---
title: "Sample Quarto Document"
author: "Dootika Vats"
date: today
format: html
editor: visual
---

## Welcome to this Sample Document

**Solving quadratic equations**

Consider equations of the form $ax^{2}+bx+c=0$. This is a quadratic equation; changing the values of $a$, $b$, and $c$, changes the shape of the quadratic. There are two solutions to a quadratic equation:

$$x=\frac{{-b \pm \sqrt{b^{2} - 4ac }}}{2a}
$$

Let's draw a quadratic curve for different values of $a$, $b$ and $c$.

```{r}
a <- 10
b <- 3
c <- -5
x <- seq(-2, 2, length = 1e3)
fx <- a*x^2 + b*x + c
plot(x, fx, type = 'l', xlab = "x", ylab = "Quadratic function")
#adding line for solution
abline(h = 0, lty = 2, col = "red")
```

In the above, we can also change the way the plot appears. I will return the above code, without displaying the code, and only producing the plot of figure height "4" units and figure width "4" units

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
